### 2020-05-26 writeup/proof_taskonomy.tex

\section{Extension to the Transfer Function of Taskonomy}

We study the transfer function of taskonomy \cite{ZSSGM18}.
The algorithm is as follows.
First, we obtain the single-task estimator $\hat{\beta}_i$ from every task, for $1\le i \le t$.
This forms the shared representation $B = [\hat{\beta}_1,\hat{\beta}_2,\dots,\hat{\beta}_{t-1}]$.
Then, we learn the output layer $W_t$ on the target task by solving equation:
\begin{align}
	\bignorm{X_t B W_t - Y_t}^2
\end{align}
After getting $W_t$, we use $B W_t$ as estimator and $\te(B W_t)$ is the test error of the estimator.
We use our tools to compare $\te(B W_t)$ to $\te(\beta_t^{\STL})$.

>>>>

\section{Supplementary Materials for Section \ref{sec_taskonomy}}\label{app_proof_sec4}

### 2020-05-26 writeup/transfer.tex

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Task model similarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Data size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
	\end{subfigure}
	\caption{Comparing the test error of multi-task learning to single-task learning: (a) positive to negative transfer as the task model dissimilarity increases (cf. Proposition \ref{prop_dist_transition}) (b) positive to negative transfer as the source task data size increases (cf. Proposition \ref{prop_data_size}); (c) }
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

\begin{proposition}\label{prop_covariate}
	In the setting of Example \ref{ex_basic}, assume that $c_1 > 3$ and $c_2>3$,
and $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$. %As a parameter of $M \in \cS_{\mu}$, 
We have that $(\te(\hat{\beta}_t^{\MTL}))(M)$ as a function of $M\in \cS_{\mu}$ satisfies that 
\be\label{formular_covariate0} (\te(\hat{\beta}_t^{\MTL}))(\mu \id) \le \left[1+ C\left(\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}\right)\right]\cdot\min_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M) ,\ee
where the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $C_1$, but otherwise does not depend on $c_1$ and $c_2$. 
	%is minimized when $M = \mu\id$.
\end{proposition}

Proposition \ref{prop_covariate} shows that when $n_1\gg n_2$, $te(\hat{\beta}_t^{\MTL})$ is minimized when $\Sigma_1$ and $\Sigma_2$ are proportional to each other.
In other words, there is no covariate shift between the source task data and target task data.
This provides evidence that covariate shift is unfavorable when there are many source task datapoints,
To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.

\begin{example}\label{ex_covariate}
	Within the setting of Example \ref{ex_cov_family}, we compare two cases: (i) when $M = \id$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	For simplicity, we assume that $d = 0$.
	Hence the two tasks have the same model parameters.
%	now consider another case when $\Sigma_1$ and $\Sigma_2$ have complementary eigenspaces. Suppose $\Sigma_1$ and $\Sigma_2$ have the eigendecomposition
%$$\Sigma_1^{1/2} = 1+ U\Lambda U^\top, \quad \Sigma_2^{1/2} = 1+ V\Lambda V,$$
%where
%$$\Lambda = \diag(\wt\lambda_1,\cdots, \wt\lambda_{p/2}), \quad U= (u_1,\cdots, u_{p/2}), \quad V= (v_1,\cdots, v_{p/2}).$$
%If $V=U_\perp$, i.e. the vectors $v_1,\cdots, v_{p/2}$ are perpendicular to the vectors $u_1,\cdots, u_{p/2}$, then
%$$M=\Sigma_1^{1/2} \Sigma_2^{-1/2}=(1+\Lambda)UU^\top + (1+\Lambda)^{-1}V V^\top .$$
%	As a concrete example we consider the case where $\wt \lambda_1=\cdots=\wt\lambda_{p/2}$ and we denote $\lambda:=1+\wt\lambda_1$. Thus for $M$, the first $p/2$ singular values are equal to $\lambda$ and the rest ones are equal to $\lambda^{-1}$.

	In Figure \ref{fig_te_complement}, we plot the test error of the target task for $n_2 = 4p$ and $n_1$ ranging from $p$ to $20p$.
	Second, we observe the following two phases as we increase $n_1 / p$.
	\begin{itemize}
		\item When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
		\item When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
	\end{itemize}
\end{example}

A theoretical justification of Example \ref{ex_covariate} can be found in Appendix \ref{app_covariate}.

@@

While our focus is on multi-task learning, we show that our results extend to transfer learning settings naturally.
We provide an analysis of the transfer function of Taskonomy (Figure 4 in \cite{ZSSGM18}) using our setup.
Specifically, the source task encoder consists of the representations learnt from one or more source tasks.
The transfer function then tries to fix the target task data to the source task encoder.

>>>>

\begin{figure}
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/model_shift_phase_transition.eps}
		\caption{Task model similarity}
		\label{fig_model_shift}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/datapoints_phase_transition.eps}
		\caption{Data size}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.32\textwidth}
		\centering
		\includegraphics[width=0.98\textwidth]{figures/complementary.eps}
		\caption{Covariate shift}
		\label{fig_covariate}
	\end{subfigure}
	\caption{Comparing the test error of multi-task learning to single-task learning: (a) positive to negative transfer as the task model dissimilarity increases (cf. Proposition \ref{prop_dist_transition}) (b) positive to negative transfer as the source task data size increases (cf. Proposition \ref{prop_data_size}); (c) }
	\label{fig_model_shift_phasetrans}
\end{figure}

@@

\begin{proposition}\label{prop_covariate}
	In the setting of Example \ref{ex_basic}, assume that $\rho_1 > 3$ and $\rho_2>3$,
and $\|\Sigma_1\|\le C_1$ for some constant $C_1>0$. %As a parameter of $M \in \cS_{\mu}$, 
The test error $\te(\hat{\beta}_t^{\MTL})$, as a function of $M\in \cS_{\mu}$, is minimized at the identity matrix $\mu\id$ approximately by
$1+ \bigo{\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}}$.
%\be\label{formular_covariate0} (\te(\hat{\beta}_t^{\MTL}))(\mu \id) \le \left[1+ C\left(\frac{\rho_2}{\rho_1} + \frac{1}{\sqrt {\rho_1}}\right)\right]\cdot\min_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M) ,\ee
%where the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $C_1$, but otherwise does not depend on $c_1$ and $c_2$.
	%is minimized when $M = \mu\id$.
\end{proposition}
Proposition \ref{prop_covariate} shows that when $n_1\gg n_2$, $te(\hat{\beta}_t^{\MTL})$ is minimized when $\Sigma_1$ and $\Sigma_2$ are proportional to each other.
In other words, there is no covariate shift between the source task data and target task data.
This provides evidence that covariate shift is unfavorable when there are many source task datapoints,
To complement the result, we show an example when the statement is not true if $n_1 \le n_2$.

\begin{example}\label{ex_complement}
	Within the setting of Example \ref{ex_cov_family}, we compare two cases: (i) when $M = \id$; (ii) when $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$.
	For simplicity, we assume that $d = 0$.
	Hence the two tasks have the same model parameters.

	In Figure \ref{fig_covariate}, we plot the test error of the target task for $n_2 = 4p$ and $n_1$ ranging from $p$ to $20p$.
	Second, we observe the following two phases as we increase $n_1 / p$.
	When $n_1 \le n_2$, having complementary covariance matrices leads to lower test error compared to the case when $\Sigma_1 = \Sigma_2$.
	When $n_1 > n_2$, having complementary covariance matrices leads to higher test error compared to the case when $\Sigma_1 = \Sigma_2$.
	A theoretical justification can be found in Appendix \ref{app_covariate}.
\end{example}

@@

While our focus is on multi-task learning, we show that our results extend to transfer learning settings naturally.
We provide an analysis of the transfer function of Taskonomy \cite{ZSSGM18} using our setup.
Specifically, the source task encoder consists of the representations learnt from one or more source tasks.
The transfer function then tries to fit the target task data to the source task encoder.
For more details, we refer the reader to Figure 4 in Taskonomy.
